WSClean: an implementation of a fast, generic wide-field imager

### 2014-02-21 articles/2014/Article-WSClean/wsclean.tex
Small details, and adding joined polarization deconvolution

The new imager, called ``WSClean'', is written in the C++ language. It reads visibilities from CASA measurement sets and writes output images to Flexible Image Transport System (FITS) files. Several steps are multi-threaded using the threading module of the C++11 standard library. These are: reading and writing; gridding from and to different $w$-layers; performing the FFTs; and H\"ogbom Clean iterations (peak-finding and image subtraction, \citealt{hogbom-clean}). For the latter, intrinsics are used as well. Because of these optimizations, minor Cleaning iterations with an image size of $3072\times3072$ are performed at a rate of hundreds per second. This is fast enough to make the Clark Clean optimization \citep{clark-clean}, which consists of considering only a subset of pixels with a reduced PSF, less relevant. However, the PSF varies for imaging with non-zero $w$-values, and subtracting a constant PSF in image space leads to inaccuracies. After a number of minor iterations, it is therefore beneficial to invert the model back to the visibilities via prediction, and subtract the model directly from the visibilities \citep{wprojection-cornwell}. This is similar to the Cotton-Schwab cleaning method \citep{cotton-schwab-clean}.

@@

The used software version of CASA is ``stable release 42.0, revision 26465'', which was released September 2013. For WSClean, the latest version as of February 2014 was used. The tests were run on a high-end desktop with 32 GB of memory and a 3.20-GHz Intel Core i7-3930K processor with 6 cores that can perform 138 giga-floating point operations per second (GFLOPS). The data are stored on a multi-disk array with 5 spinning hard disks, which has a combined read rate of about 450 MB/s.
\begin{table}%
\caption{Parameter values used during benchmarks, unless otherwise mentioned.} \label{tbl:default-parameters}%
\begin{center}\begin{tabular}{rl}%
\hline
Array & MWA 128T \\
Image size & $3072 \times 3072$ \\
Angular resolution & $0.72'$ \\
Number of visibilities & $3.5 \times 10^8$ \\
Time resolution & 2~s \\
Frequency resolution & 40~kHz \\
Observation duration & 112 s\\
Bandwidth & 30.72 MHz (768 channels)\\
Central frequency & 182 MHz \\
Zenith angle at phase centre & 10\degree \\
Max $w$-value for phase centre & 172 $\lambda$ (283 m) \\
Nr. polarizations in set & 4 \\
Imaged polarization & $pp$ ($\sim$XX) \\
Imaging mode & Multi-frequency synthesis \\
Weighting & uniform \\
Data size & 18 GB \\
\hline
\end{tabular}\end{center}\end{table}

@@

Function $t_\textrm{snapshot}$ can be minimized to find the optimal snapshot duration. For WSClean with MWA parameters, we find that this function decreases but no minimum is reached for $\Delta \tau_\textrm{snapshot}<24$~h. Therefore, from a performance perspective, the snapshot duration should be as large as possible. For $w$-projection in CASA we find an optimal snapshot duration of $\sim2$ min for MWA observations.

>>>>

The new imager, called ``WSClean'', is written in the C++ language. It reads visibilities from CASA measurement sets and writes output images to Flexible Image Transport System (FITS) files. Several steps are multi-threaded using the threading module of the C++11 standard library. These are: reading and writing; gridding from and to different $w$-layers; performing the FFTs; and H\"ogbom Clean iterations (peak-finding and image subtraction, \citealt{hogbom-clean}). For the latter, intrinsics are used as well. Because of these optimizations, minor Cleaning iterations with an image size of $3072\times3072$ are performed at a rate of hundreds per second. This is fast enough to make the Clark Clean optimization \citep{clark-clean}, which consists of considering only a subset of pixels with a reduced PSF, less relevant. However, the PSF varies for imaging with non-zero $w$-values, and subtracting a constant PSF in image space leads to inaccuracies. After a number of minor iterations, it is therefore beneficial to invert the model back to the visibilities via prediction, and subtract the model directly from the visibilities \citep{wprojection-cornwell}. This is similar to the Cotton-Schwab cleaning method \citep{cotton-schwab-clean}. WSClean allows cleaning individual polarizations, or can jointly deconvolve the polarizations. In the latter case, peaks are searched in $pp^2+2pq \overline{(pq)} + qq^2$ space, where $\overline{pq}=qp$ is the complex conjugate of $pq$, and the PSF is subtracted from the polarizations with different factors.

@@

The used software version of CASA is ``stable release 42.0, revision 26465'', which was released September 2013. For WSClean, the latest version as of February 2014 was used. The tests were run on a high-end desktop with 32 GB of memory and a 3.20-GHz Intel Core i7-3930K processor with 6 cores that can perform 138 giga-floating point operations per second (GFLOPS). The data are stored on a multi-disk array with 5 spinning hard disks, which has a combined read rate of about 450 MB/s.
\begin{table}%
\caption{Parameter values used during benchmarks, unless otherwise mentioned.} \label{tbl:default-parameters}%
\begin{center}\begin{tabular}{rl}%
\hline
Array & MWA 128T \\
Image size & $3072 \times 3072$ \\
Angular pixel size & $0.72'$ \\
Number of visibilities & $3.5 \times 10^8$ \\
Time resolution & 2~s \\
Frequency resolution & 40~kHz \\
Observation duration & 112 s\\
Bandwidth & 30.72 MHz (768 channels)\\
Central frequency & 182 MHz \\
Zenith angle at phase centre & 10\degree \\
Max $w$-value for phase centre & 172 $\lambda$ (283 m) \\
Nr. polarizations in set & 4 \\
Imaged polarization & $pp$ ($\sim$XX) \\
Imaging mode & Multi-frequency synthesis \\
Weighting & uniform \\
Data size & 18 GB \\
\hline
\end{tabular}\end{center}\end{table}

@@

Function $t_\textrm{snapshot}$ can be minimized to find the optimal snapshot duration. For WSClean with MWA parameters, we find that this function decreases but no minimum is reached for $\Delta \tau_\textrm{snapshot}<24$~h. Therefore, from a performance perspective, the snapshot duration should be as large as possible. In practise, the beam needs to be corrected on small timescales. A snapshot duration of more than a few minutes is therefore not possible. For $w$-projection in CASA we find an optimal snapshot duration of $\sim2$ min for MWA observations.
